On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-07-09 manuscripts/PNAS/chazthm_pnas.tex
tiny edit
newness

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Charles J. Garfinkle}
\author[a,1]{Christopher J. Hillar}

\affil[a]{Redwood Center for Theoretical Neuroscience, Berkeley, CA, USA}

@@

@@

All other proofs are relegated to the Supplemental. 
Finally, we present several applications in Discussion Sec.~\ref{Discussion}.

@@

\begin{definition}\label{sip}
Given $E \subseteq 2^{[m]}$, the \textbf{star} $F(i)$ at $i$ is the set of $S \in E$ with $i \in S$. We say $E$ has the \textbf{singleton intersection property} (\textbf{SIP}) when $\cap F(i) = \{i\}$ for all $i \in [m]$. 
\end{definition}

It is easy to verify that for every $k < m$, there is a regular $k$-uniform hypergraph that satisfies the SIP; for instance, consecutive intervals of length $k$ in some cyclic order on $[m]$.

Next, we describe a quantitative version of the spark condition.  
The \emph{lower bound} \cite{Grcar10} of a matrix $\mathbf{M} \in \mathbb R^{n \times m}$ is the largest number $\alpha$ such that \mbox{$\|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2$} for all $\mathbf{x} \in \mathbb{R}^m$. By compactness of the unit sphere, every injective linear map has a nonzero lower bound; hence, if $\mathbf{M}$ satisfies \eqref{SparkCondition}, then each submatrix formed from $2k$ of its columns or less has a nonzero lower bound. This motivates the following definition: 
\begin{align*}
L_k(\mathbf{M}) := \frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
\end{align*}

@@

\begin{theorem}\label{DeterministicUniquenessTheorem}
Let $\mathbf{A}~\in~\mathbb{R}^{n \times m}$ satisfy \eqref{SparkCondition} for $k < m$, and suppose for some $\bar m$ that more than \mbox{$(k-1){\bar m \choose k}$} $k$-sparse \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} are supported in general linear position on each set in an $\ell$-regular $E \subseteq {[m] \choose k}$ with the SIP.  There is a constant\footnote{We delay defining the constant $C_1$ until Section \ref{DUT} (\eqref{Cdef1}).} $C_1 > 0$ for which the following holds for $\varepsilon < L_2(\mathbf{A}) / C_1$:

Every matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\bar x}_1, \ldots, \mathbf{\bar x}_N \in \mathbb{R}^{\bar m}$ with  \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\bar x}_i\|_2 \leq \varepsilon$} for all $i$ must have $\bar m \geq m$ and, provided $p = \bar m - \ell(\bar m - m)$ is positive, necessarily satisfy:
\begin{align}\label{Cstable}
\|(\mathbf{A}_J- \mathbf{B}_{\bar J} \mathbf{PD} )_j\|_2 \leq C_1 \varepsilon, \ \  \text{for $j \in J$},
\end{align}
for some $J \in {[m] \choose p}$, $\bar J \in {[\bar m] \choose p}$, permutation matrix $\mathbf{P}$, and invertible diagonal matrix $\mathbf{D}$.

Furthermore, if $\varepsilon < L_{2k}(\mathbf{A}_J) / C_1$, then $\mathbf{B}_{\bar J}$ also satisfies \eqref{SparkCondition} with $L_{2k}(\mathbf{B}_{\bar J}\mathbf{PD}) \geq L_{2k}(\mathbf{A}_J) - C_1 \varepsilon$, and:
\begin{align}\label{b-PDa}
\|\mathbf{x}_i^J - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\bar x}_i^{\bar J}\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}_i^{J}\|_1 }{ L_{2k}(\mathbf{A}_J) -  C_1\varepsilon } \right) \varepsilon,\ \ \   \text{for $i \in [N]$}.
\end{align}
\end{theorem}

To be clear, Thm.~\ref{DeterministicUniquenessTheorem} says that the smaller the difference $\bar m - m$, the more columns and coefficients of the original dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$ contained (up to noise) in the appropriately scaled dictionary $\mathbf{B}$ and codes $\mathbf{\bar x}_i$. 
The implication in the case $\bar{m} = m$ is precisely that $Y = \{\mathbf{Ax}_1, \ldots, \mathbf{Ax}_N\}$ has a stable $k$-sparse representation in $\mathbb{R}^m$, with \eqref{def1} guaranteed provided $\varepsilon$ in Def.~\ref{maindef} does not exceed: 
\begin{align}\label{epsdel}
\varepsilon(\delta_1, \delta_2) := \min \left\{ \frac{\delta_1}{ C_1 }, \frac{ \delta_2 L_{2k}(\mathbf{A})}{ 1 + C_1 \left( \max_{i \in [N]} \|\mathbf{x}_i\|_1  + \delta_2 \right) } \right\}.
\end{align}

It so happens that producing sparse codes in general linear position is straightforward with a ``Vandermonde'' matrix construction (e.g., see \cite{Hillar15}). We therefore have:

@@

As already mentioned, it is easy to construct a regular hypergraph $E \subseteq {[m] \choose k}$ with the SIP having cardinality $|E| = m$, implying the lower bound for sample size $N$ from the introduction. In many cases, however, the SIP can be achieved with far fewer supports; for example, when $k = \sqrt{m}$ one can take $E$ to be the $2k$ rows and columns formed by arranging $[m]$ in a square grid.

@@

@@

Since $f$ is an analytic function, having one substitution of real numbers with $f(\mathbf{A}) \neq 0$ implies that its zeroes form a set of measure zero. Fortunately, such a matrix $\mathbf{A}$ is easily constructed by adding rows of zeroes to any $\min(2k,m) \times m$ Vandermonde matrix $[\gamma_i^j]_{i,j=1}^{k,m}$ with distinct $\gamma_i$ (so that each term in the product above is nonzero).
Hence, almost every real $n \times m$ matrix $\mathbf{A}$ with $n \geq \min(2k,m)$ satisfies \eqref{SparkCondition}.

@@

% ======== b - PDa =========
First, note that $\|\mathbf{x}\|_1 \leq \sqrt{k} \|\mathbf{x}\|_2$ for $k$-sparse $\mathbf{x} \in \mathbb{R}^m$, which by definition of $L_k$ implies the following inequality:
\begin{align}\label{delrho}
L_k(\mathbf{A}) \leq \frac{\|\mathbf{A}\mathbf{x}\|_2}{\sqrt{k} \|\mathbf{x}\|_2} 
\leq  \max_{j \in [m]}\|\mathbf{A}_j\|_2,
\end{align}
which we frequently use.

We begin by showing how dictionary recovery \eqref{Cstable} already implies sparse code recovery \eqref{b-PDa} when \mbox{$\varepsilon < L_{2k}(\mathbf{A}) / C_1$}, assuming without loss of generality that $\bar m = m$. For $2k$-sparse $\mathbf{x} \in \mathbb{R}^m$, the triangle inequality gives \mbox{$\|(\mathbf{A}-\mathbf{BPD})\mathbf{x}\|_2  \leq C_1\varepsilon \|\mathbf{x}\|_1 \leq C_1 \varepsilon \sqrt{2k}  \|\mathbf{x}\|_2$}. Thus, 
\begin{align*}
\|\mathbf{BPD}\mathbf{x}\|_2 
&\geq | \|\mathbf{A}\mathbf{x}\|_2 - \|(\mathbf{A}-\mathbf{BPD})\mathbf{x}\|_2 | \\
&\geq \sqrt{2k} (L_{2k}(\mathbf{A}) -  C_1\varepsilon) \|\mathbf{x}\|_2.
\end{align*}

@@

\begin{lemma}\label{MainLemma}
Suppose $\mathbf{A} \in \mathbb{R}^{n \times m}$ satisfies \eqref{SparkCondition} for some $k < m$ and that $E \subseteq {[m] \choose k}$ is $\ell$-regular with the SIP. There exists $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$.

If for some  $\mathbf{B} \in \mathbb{R}^{n \times \bar m}$ and map $\pi: E \mapsto {[\bar m] \choose k}$ we have:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon, \ \  \text{for $S \in E$},
\end{align}
then $\bar m \geq m$, and if $p = \bar m - \ell(\bar m - m)$ is positive:
\begin{align}\label{MainLemmaBPD}
\|(\mathbf{A}_J - \mathbf{B}_{\bar J}\mathbf{PD})_j\|_2 \leq C_2 \varepsilon, \ \  \text{for } j \in J,
\end{align}
for some $J \in {[m] \choose p}$, $\bar J \in {[\bar m] \choose p}$, permutation matrix $\mathbf{P}$, and invertible diagonal matrix $\mathbf{D}$.
\end{lemma}

@@

Since there are more than $(k-1){\bar m \choose k}$ vectors $\mathbf{x}_i$ supported on $S$, by the pigeonhole principle there must be some $\bar S \in {[\bar m] \choose k}$ and set of $k$ indices $K$ such that the supports of all $\mathbf{x}_i$ and $\mathbf{\bar x}_i$ with $i \in K$ are contained in $S$ and $\bar S$, respectively.

It follows from the general linear position of the $\mathbf{x}_i$ and the linear independence of every $2k$ columns of $\mathbf{A}$ that $L_k(\mathbf{AX}_{K}) > 0$; that is, the columns of the $n \times k$ matrix $\mathbf{AX}_K$ form a basis for $\text{\rm Span}\{\mathbf{A}_{S}\}$. Fixing $\mathbf{0} \neq \mathbf{y} \in \text{\rm Span}\{\mathbf{A}_{S}\}$, there exists $\mathbf{0} \neq \mathbf{c} = (c_1, \ldots, c_k) \in \mathbb{R}^k$ such that $\mathbf{y} = \mathbf{AX}_K\mathbf{c}$. Setting \mbox{$\mathbf{\bar{y}} = \mathbf{B\bar{X}}_K\mathbf{c} \in \text{\rm Span}\{\mathbf{B}_{\bar S}\}$}, we have:
\begin{align*}
\|\mathbf{y} - \mathbf{\bar{y}}\|_2 
&= \|\sum_{i=1}^k c_i(\mathbf{AX}_K - \mathbf{B\bar{X}}_K)_i\|_2
\leq \varepsilon \sum_{i=1}^k |c_i| \\
&\leq \varepsilon \sqrt{k}  \|\mathbf{c}\|_2 
\leq \frac{\varepsilon}{L_k(\mathbf{AX}_K)} \|\mathbf{y}\|_2,
\end{align*}
where the last inequality follows directly from the definition of $L_k$. Finally, from Def.~\ref{dDef} we have:
\begin{align}\label{rhs222}
d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{\bar S}\}) 
\leq \frac{\varepsilon}{  L_k(\mathbf{AX}_{K}) },
\end{align}
%
from which the result follows by \eqref{Cdef1} and Lem.~\ref{MainLemma}.
\end{proof}

@@

\section{Discussion}\label{Discussion}

In this note, we generalized the approach of \cite{Hillar15} to prove the stability of unique solutions to Probs.~\ref{InverseProblem} and \ref{OptimizationProblem} while significantly reducing the known sample complexity. 
Our results justify the application of the sparse linear coding model to blind source separation problems, wherein the goal is to infer the generating dictionary and sparse codes from noisy measurements, \eqref{LinearModel}. We also collect a set of useful mathematical tools and basic facts for future research.
The main motivation for this work, however, was to understand how seemingly universal representations emerge from sparse coding models fit to natural data by a variety of methodologies. We elaborate on these applications below after discussing theoretical aspects.

What we have shown here is that the sparse linear coding model generally produces a \textit{well-posed} inverse problem to be approximately solved by a numerical algorithm. This early concept of Hadamard \cite{Hadamard1902} can be paraphrased as the idea that inferences from observations should be robust to the inevitable uncertainty in measurement. In other words, a small perturbation of the data should result in only slightly different inferred parameters. In this regard, for the sparse linear coding model we demonstrate a linear relationship, \eqref{epsdel}, between measurement noise and the deviation of any solution from the true parameters, with explicit constants expressed in terms of these parameters. Moreover, we show that even if the meta-parameter for the number of dictionary elements is overestimated, a subset of parameters may still be identifiable up to noise.
It would therefore be of practical utility to determine the best possible dependence of $\varepsilon$ on $\delta_1, \delta_2$ in Def.\ref{maindef} as well as the minimal requirements on the number and diversity of generating codes, and we hope that other researchers continue to improve and extend our results. We remark that our constants have been derived for deterministic ``worst-case'' noise, whereas the ``effective'' noise might be smaller when sampled from a distribution; in such cases, the constants 
will improve as well.

One notable component of our contribution is a combinatorial criteria (regular hypergraphs satisfying the singleton intersection property, Def.~\ref{sip}) for the support sets of sparse codes key to the identification of the dictionary. Fully understanding those combinatorial designs allowing for stable sparse representations is an interesting research area for the future. For instance, whether there is a support set size $N$ that is polynomial in $m,k$ gauranteeing the conclusions of Thm.~\ref{DeterministicUniquenessTheorem} has implications for the computational complexity of Probs.~\ref{InverseProblem} and \ref{OptimizationProblem} and related questions \cite{Tillmann15}.

>>>>

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,b, 1]{Charles J. Garfinkle}
\author[a,1]{Christopher J. Hillar}

\affil[a]{Redwood Center for Theoretical Neuroscience, Berkeley, CA, USA}
\affil[b]{Helen Wills Neuroscience Institute, UC Berkeley}

@@

\pagebreak

@@

All other proofs are relegated to the Section \ref{proofs}. 
Finally, we present several applications in Discussion Sec.~\ref{Discussion}.

@@

\begin{definition}\label{sip}
Given $E \subseteq 2^{[m]}$, the \textbf{star} $H(i)$ at $i$ is the set of $S \in E$ with $i \in S$. We say $E$ has the \textbf{singleton intersection property} (\textbf{SIP}) when $\cap H(i) = \{i\}$ for all $i \in [m]$. %We say $E$ has the SIP \textbf{of order $p$} when $\cap F(i) = \{i\}$ for every node $i$ for which $\text{deg(i)} \geq d_p$, where $d_1 \geq \ldots \geq d_m$ is the degree sequence of $E$.
\end{definition}

Next, we describe a quantitative version of the spark condition.  
The \emph{lower bound} $L$ of a matrix $\mathbf{M} \in \mathbb R^{n \times m}$ is the largest number $\alpha$ such that \mbox{$\|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2$} for all $\mathbf{x} \in \mathbb{R}^m$ \cite{Grcar10}. By compactness of the unit sphere, every injective linear map has a nonzero lower bound; hence, if $\mathbf{M}$ satisfies \eqref{SparkCondition}, then each submatrix formed from $2k$ of its columns or less has a nonzero lower bound. This motivates the following definition: 
%\begin{align*}
%L_k(\mathbf{M}) := \frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
%\end{align*}
\begin{align*}
L_E(\mathbf{M}) := \inf \left\{ \frac{ \|\mathbf{M}_{S \cup S'}\mathbf{x}\|_2 }{ \sqrt{2k} \|\mathbf{x}\|_2} : \mathbf{x} \in \mathbb{R}^{|S\cup S'|}, \ \ S,S' \in E \right\}
%\frac{1}{\sqrt{k}}\max \{ \alpha : \|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2 \text{ for all $k$-sparse } \mathbf{x}\}.
\end{align*} 
%
and we write $L_{2k}:= L_E$ when $E = {[m] \choose k}$. Note that $ L = L_{\{[m]\}}$ whereas $L_{[m]} = L_2$. [TODO: note that $L_2$ implied by $L_E$]

@@

\begin{theorem}\label{DeterministicUniquenessTheorem}
Fix a matrix $\mathbf{A} \in \mathbb{R}^{n \times m}$ with $L_E(\mathbf{A}) > 0$ for some regular $E \subseteq {[m] \choose k}$ with the SIP. If the sequence \mbox{$\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$} contains, for each $S \in E$, more than $(k-1){m \choose k}$ vectors in general linear position supported on $S$, then there is a constant\footnote{We delay defining the constant $C_1$ until Section \ref{DUT} (\eqref{Cdef1}).} $C_1 > 0$ for which the following holds for all $\varepsilon < L_{2}(\mathbf{A}) / C_1$:

Every real $n \times \bar m$ matrix $\mathbf{B}$ for which there exist $k$-sparse $\mathbf{\bar x}_1, \ldots, \mathbf{\bar x}_N$ satisfying \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\bar x}_i\|_2 \leq \varepsilon$} for all $i$ has $\bar m \geq m$ and, provided $\bar m < mr/(r-1)$, an associated permutation $\mathbf{P}$ and invertible diagonal $\mathbf{D}$ such that:
\begin{align}\label{Cstable}
\|(\mathbf{A_J})_j- \mathbf{B}_{\bar J} \mathbf{PD}_j\|_2 \leq C_1 \varepsilon \ \ \text{for all } j \in J,
\end{align}
%
for some non-empty $J, \bar J \subseteq [m]$ of size $\bar m - r(\bar m - m)$.

Moreover, if in fact $L_{2k}(\mathbf{A}) > 0$ and $\varepsilon < L_{2k}(\mathbf{A}) / C_1$ then $L_{2k}(\mathbf{B}\mathbf{PD}) \geq L_{2k}(\mathbf{A}) - C_1 \varepsilon$ and:
\begin{align}\label{b-PDa}
\|\mathbf{x}^J_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\bar x}^{\bar J}_i\|_1 &\leq  \left( \frac{ 1+C_1 \|\mathbf{x}^{J}_i\|_1 }{ L_{2k}(\mathbf{A}) -  C_1\varepsilon } \right) \varepsilon,\ \ \   \text{for $i \in [N]$}.
\end{align}
\end{theorem}
% You can recover matrices that don't satisfy the spark condition, but to uniquely recover k-sparse codes (since no limitation on their supports!) you need to have the spark condition. i guess...? Or can we use our knowledge of the particular permutation P to get rid of spark condition?

%To be clear, the implication of Thm.~\ref{DeterministicUniquenessTheorem} is that $Y = \{\mathbf{Ax}_1, \ldots, \mathbf{Ax}_N\}$ has a stable $k$-sparse representation in $\mathbb{R}^m$, with \eqref{def1} guaranteed provided $\varepsilon$ in Def.~\ref{maindef} does not exceed: 
To be clear, Thm.~\ref{DeterministicUniquenessTheorem} says that the smaller the difference $\bar m - m$, the more columns and coefficients of the original dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$ contained (up to noise) in the appropriately scaled dictionary $\mathbf{B}$ and codes $\mathbf{\bar x}_i$. The implication in the case $\bar m = m$ is that $Y = \{\mathbf{Ax}_1, \ldots, \mathbf{Ax}_N\}$ has a stable $k$-sparse representation in $\mathbb{R}^m$, with \eqref{def1} guaranteed provided $\varepsilon$ in Def.~\ref{maindef} does not exceed: 
\begin{align}\label{epsdel}
\varepsilon(\delta_1, \delta_2) := \min \left\{ \frac{\delta_1}{ C_1 }, \frac{ \delta_2 L_{2k}(\mathbf{A})}{ 1 + C_1 \left( \max_{i \in [N]} \|\mathbf{x}_i\|_1  + \delta_2 \right) } \right\}.
\end{align}

%In fact, a more general result (stated clearly in the next section) can be gleaned from our method of proving Thm.~\ref{DeterministicUniquenessTheorem}. Briefly, in cases where $\mathbf{B}$ has $\bar m \neq m$ columns, or $E$ is not regular or only partially satisfying the SIP, a relation between $\bar m$ and the degree sequence of nodes in $E$ gives indices $J \subseteq [m]$ defining a submatrix $\mathbf{A}_J$ and subvectors $\mathbf{x}_i^J$ that are recoverable in the sense of \eqref{Cstable} and \eqref{b-PDa}. For example, if $E$ is $\ell$-regular with the SIP but $m \leq \bar m < m\ell/(\ell - 1)$ then we have nonzero $|J| = \bar m - \ell(\bar m - m)$. The implication here is that the smaller the difference $\bar m - m$, the more columns and code entries of the original $n \times m$ dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$ contained (up to noise) in the appropriately scaled $n \times \bar m$ dictionary $\mathbf{B}$ and codes $\mathbf{\bar x}_i$. When $\bar m = m$, we recover Thm.~\ref{DeterministicUniquenessTheorem}.
Regarding the assumptions of Thm.~\ref{DeterministicUniquenessTheorem}, it is easy to verify that for every $k < m$, there is a regular $k$-uniform hypergraph that satisfies the SIP; for instance, consecutive intervals of length $k$ in some cyclic order on $[m]$.
In fact, even if $E$ is not regular or only partially satisfies the SIP, a relation between $\bar m$ and the degree sequence of nodes in $E$ may give the indices $J \subseteq [m]$. For sake of brevity, we delay to the next section a clear statement of this more general result.

It also happens that producing sparse codes $\mathbf{x}_i$ in general linear position is straightforward with a ``Vandermonde'' matrix construction (e.g., see \cite{Hillar15}). We therefore have:

@@

As already mentioned, there exist $k$-uniform regular hypergraphs $E$ with the SIP having cardinality $|E| = m$, implying the lower bound for sample size $N$ from the introduction. In many cases, however, the SIP can be achieved with far fewer supports; for example, when $k = \sqrt{m}$, take $E$ to be the $2k$ rows and columns formed by arranging $[m]$ in a square grid.

@@

%\pagebreak

@@

Since $f$ is an analytic function, if there exists one substitution of a real matrix $\mathbf{A}$ such that $f(\mathbf{A}) \neq 0$ then the zeroes of $f$ in fact form a set of measure zero. Fortunately, such a matrix $\mathbf{A}$ is easily constructed by adding rows of zeroes to any $\min(2k,m) \times m$ Vandermonde matrix $[\gamma_i^j]_{i,j=1}^{k,m}$ with distinct $\gamma_i$ (so that each term in the product above is nonzero). Hence, almost every real $n \times m$ matrix with $n \geq \min(2k,m)$ satisfies \eqref{SparkCondition}.

@@

As mentioned in the previous section, Thm.~\ref{DeterministicUniquenessTheorem} is a particular case of a more general result that forgoes the assumptions of regularity and SIP on the hypergraph $E \subseteq {[m] \choose k}$. If instead we require only that the stars $\cap H(i)$ intersect at singletons for all $i \leq q$ (assuming that the nodes of $E$ are labeled in some order of non-increasing degree), we have that $\bar m \geq k|E| / \deg(1)$ and, provided $\bar m < k|E| / (\deg(1) - 1)$, the non-empty submatrix $J$ is of size equal to the largest number $p$ satisfying:
\begin{align}\label{pcond}
\sum_{i=\ell}^{m} \deg(i) > (\bar m + 1 - \ell) (\deg(\ell) - 1) \ \ \text{for all } \ell \leq p \leq q.
\end{align}
Specifically, $J$ consists of the union of the set of all nodes of degree exceeding $\deg(p)$ and some subset of those nodes with degrees equal to $\deg(p)$. For the benefit of the reader, we do not prove this more general result below; it can be gleaned by examining how exactly Lemma \ref{NonEmptyLemma} is incorporated into the proof of Lem.~\ref{MainLemma}.
%As mentioned in the previous section, Thm.~\ref{DeterministicUniquenessTheorem} is a particular case of a more general result, which requires a looser set of constraints on the hypergraph $E$ and which applies to $n \times \bar m$ matrices $\mathbf{B}$ (and $\bar m$-dimensional codes $\mathbf{\bar x}_i$) with $\bar m \neq m$:

%Fix $\bar m$ and suppose the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} hold, only now with the constraints on $E \subseteq {[m] \choose k}$ being just that $|\cap H(i)| = 1$ for all $i \leq q$ (assuming w.l.o.g. that the nodes of $E$ are labeled in some order of non-increasing degree), and with more than $(k-1){\bar m \choose k}$ vectors $\mathbf{x}_i$ supported in g.l.p. on each $S \in E$. Then we must have $\bar m \geq k|E| / \deg(1)$ and, provided $\bar m < k|E| / (\deg(1) - 1)$, the guarantee \eqref{Cstable} holds for a submatrix $\mathbf{A}_J$, where $J \subseteq[m]$ is nonempty and of a size equal to the largest number $p$ satisfying:
%\begin{align}\label{pcond}
%\sum_{i=\ell}^{m} \deg(i) > (\bar m + 1 - \ell) (\deg(\ell) - 1) \ \ \text{for all } \ell \leq p \leq q.
%\end{align}
%Specifically, $J$ consists of the union of the set of all nodes of degree exceeding $\deg(p)$ and some subset of those nodes with degrees equal to $\deg(p)$.

%For the benefit of the reader, we prove below the case where we forgo only the constraint that $\bar m = m$. This yields the implication $\bar m \geq m$ and \eqref{pcond} reduces to $|J| = \bar m - r(\bar m - m)$. The extension to the general result above can be seen by examining how exactly Lemma \ref{NonEmptyLemma} is incorporated into the overall proof.

% ======== b - PDa =========

We now begin our proof of Thm.~\ref{DeterministicUniquenessTheorem} by showing how dictionary recovery \eqref{Cstable} already implies sparse code recovery \eqref{b-PDa} when $L_{2k}(\mathbf{A}) > 0$ and \mbox{$\varepsilon < L_{2k}(\mathbf{A}) / C_1$}, assuming without loss of generality that $\bar m = m$. First, note that $\|\mathbf{x}\|_1 \leq \sqrt{k} \|\mathbf{x}\|_2$ for $k$-sparse $\mathbf{x} \in \mathbb{R}^m$, which by definition of $L_E$ implies the following frequently applied inequality valid for any $E \subseteq 2^{[m]}$:
\begin{align}\label{delrho}
L_{E}(\mathbf{A}) \leq \frac{\|\mathbf{A}\mathbf{x}\|_2}{\sqrt{2k} \|\mathbf{x}\|_2} 
\leq  \max_{j \in [m]}\|\mathbf{A}_j\|_2,
\end{align}
For $k$-sparse $\mathbf{x} \in \mathbb{R}^m$, the triangle inequality gives \mbox{$\|(\mathbf{A}-\mathbf{BPD})\mathbf{x}\|_2  \leq C_1\varepsilon \|\mathbf{x}\|_1 \leq C_1 \varepsilon \sqrt{2k}  \|\mathbf{x}\|_2$}. Thus, 
\begin{align*}
\|\mathbf{BPD}\mathbf{x}\|_2 
&\geq | \|\mathbf{A}\mathbf{x}\|_2 - \|(\mathbf{A}-\mathbf{BPD})\mathbf{x}\|_2 | \\
&\geq \sqrt{2k} (L_{2k}(\mathbf{A}) -  C_1\varepsilon) \|\mathbf{x}\|_2.
\end{align*}

@@

\begin{lemma}\label{MainLemma}
Suppose $\mathbf{A} \in \mathbb{R}^{n \times m}$ has $L_E(\mathbf{A}) > 0$ for some $r$-regular $E \subseteq {[m] \choose k}$ with the SIP. There exists $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$:

If for some  $\mathbf{B} \in \mathbb{R}^{n \times \bar m}$ and map $\pi: E \mapsto {[\bar m] \choose k}$ we have:
\begin{align}\label{GapUpperBound}
d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon, \ \  \text{for $S \in E$},
\end{align}
then $\bar m \geq m$, and provided $\bar m < mr /(r-1)$, there exists a permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that:
\begin{align}\label{MainLemmaBPD}
\|(\mathbf{A}_J)_j - \mathbf{BPD}_j\|_2 \leq C_2 \varepsilon, \ \  \text{for } j \in J,
\end{align}
for some non-empty $J$ of size $\bar m - r(\bar m - m)$.
\end{lemma}

%\begin{lemma}\label{MainLemma}
%Fix $\mathbf{A} \in \mathbb{R}^{n \times m}$ with $L_E(\mathbf{A}) > 0$ for some regular $E \subseteq {[m] \choose k}$ with the SIP. There exists a constant $C_2 > 0$ for which the following holds for all $\varepsilon < L_2(\mathbf{A}) / C_2$:

%If a matrix $\mathbf{B} \in \mathbb{R}^{n \times m}$ and map $\pi: E \mapsto {m \choose k}$ satisfy:
%\begin{align}\label{GapUpperBound}
%d(\text{\rm Span}\{\mathbf{A}_{S}\}, \text{\rm Span}\{\mathbf{B}_{\pi(S)}\}) \leq \varepsilon, \ \ \text{for $S \in E$},
%\end{align}
%then there exist a permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that:
%\begin{align}\label{MainLemmaBPD}
%\|\mathbf{A}_j - \mathbf{BPD}_j\|_2 \leq C_2 \varepsilon, \ \  \text{for } j \in [m],
%\end{align}
%\end{lemma}

@@

Since there are more than $(k-1){\bar m \choose k}$ vectors $\mathbf{x}_i$ supported on $S$, by the pigeonhole principle there must be some $\bar S \in {[\bar m] \choose k}$ and set of $k$ indices $K \subseteq I(S)$ such that the supports of all $\mathbf{\bar x}_i$ with $i \in K$ are subsets of $\bar S$.

It follows from the general linear position of the $\mathbf{x}_i$ and the linear independence of the columns of $\mathbf{A}_S$ that $L(\mathbf{AX}_{K}) > 0$; that is, the columns of the $n \times k$ matrix $\mathbf{AX}_K$ form a basis for $\text{\rm Span}\{\mathbf{A}_{S}\}$. Fixing $\mathbf{0} \neq \mathbf{y} \in \text{\rm Span}\{\mathbf{A}_{S}\}$, there then exists $\mathbf{0} \neq \mathbf{c} = (c_1, \ldots, c_k) \in \mathbb{R}^k$ such that $\mathbf{y} = \mathbf{AX}_K\mathbf{c}$. Setting \mbox{$\mathbf{\bar{y}} = \mathbf{B\bar{X}}_K\mathbf{c} \in \text{\rm Span}\{\mathbf{B}_{\bar S}\}$}, we have:
\begin{align*}
\|\mathbf{y} - \mathbf{\bar{y}}\|_2 
&= \|\sum_{i=1}^k c_i(\mathbf{AX}_K - \mathbf{B\bar{X}}_K)_i\|_2
\leq \varepsilon \sum_{i=1}^k |c_i| \\
&\leq \varepsilon \sqrt{k}  \|\mathbf{c}\|_2 
\leq \frac{\varepsilon}{L(\mathbf{AX}_K)} \|\mathbf{y}\|_2,
\end{align*}
where the last inequality follows directly from the definition of $L$. From Def.~\ref{dDef} we have:
\begin{align}\label{rhs222}
d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{\bar S}\}) 
\leq \frac{\varepsilon}{  L(\mathbf{AX}_{K}) } \leq \varepsilon \frac{C_1}{C_2},
\end{align}
%
where the second inequality follows from $L(\mathbf{AX}_{K}) \geq L_k(\mathbf{AX}_{I(S)})$ and \eqref{Cdef1}. Since $\varepsilon < L_2(\mathbf{A})/C_1$, the result follows by Lem.~\ref{MainLemma}, .
\end{proof}

@@

\section{Discussion}\label{Discussion}
[ TODO ]
%[mention dictionary size independence]
%In this note, we generalized the approach of \cite{Hillar15} to prove the stability of unique solutions to Probs.~\ref{InverseProblem} and \ref{OptimizationProblem} while significantly reducing the known sample complexity. Our results justify the application of the sparse linear coding model to blind source separation problems, wherein the goal is to infer the generating dictionary and sparse codes from noisy measurements, \eqref{LinearModel}. We also collect a set of useful mathematical tools and basic facts for future research. The main motivation for this work, however, was to understand how seemingly universal representations emerge from sparse coding models fit to natural data by a variety of methodologies. We elaborate on these applications below after discussing theoretical aspects.

%What we have shown here is that the sparse linear coding model generally produces a \textit{well-posed} inverse problem to be approximately solved by a numerical algorithm. This early concept of Hadamard \cite{Hadamard1902} can be paraphrased as the idea that inferences from observations should be robust to the inevitable uncertainty in measurement. In other words, a small perturbation of the data should result in only slightly different inferred parameters. In this regard, for the sparse linear coding model we demonstrate a linear relationship, \eqref{epsdel}, between measurement noise and the deviation of any solution from the true parameters, with explicit constants expressed in terms of these parameters. Moreover, we show that even if the meta-parameter for the number of dictionary elements is overestimated, a subset of parameters may still be identifiable up to noise. It would therefore be of practical utility to determine the best possible dependence of $\varepsilon$ on $\delta_1, \delta_2$ in Def.\ref{maindef} as well as the minimal requirements on the number and diversity of generating codes, and we hope that other researchers continue to improve and extend our results. We remark that our constants have been derived for deterministic ``worst-case'' noise, whereas the ``effective'' noise might be smaller when sampled from a distribution; in such cases, the constants will improve as well.

%One notable component of our contribution is a combinatorial criteria (regular hypergraphs satisfying the singleton intersection property, Def.~\ref{sip}) for the support sets of sparse codes key to the identification of the dictionary. Fully understanding those combinatorial designs allowing for stable sparse representations is an interesting research area for the future. For instance, whether there is a support set size $N$ that is polynomial in $m,k$ gauranteeing the conclusions of Thm.~\ref{DeterministicUniquenessTheorem} has implications for the computational complexity of Probs.~\ref{InverseProblem} and \ref{OptimizationProblem} and related questions \cite{Tillmann15}.

%A technical difficulty in proving Thm.~\ref{DeterministicUniquenessTheorem} was the absence of a spark condition assumption on solutions to Prob.~\ref{InverseProblem}. Although mathematically interesting that no such requirement is necessary, there are other reasons to seek out such a theoretical guarantee. For instance, it is difficult to ensure that an algorithm maintain a dictionary satisfying \eqref{SparkCondition} at each iteration; indeed, even certifying a dictionary has this property is likely intractable given its NP-hardness \cite{tillmann2014computational}.
